Equations & R Code

### 2017-02-20 manuscript/coexistence-stability_ms.Rmd
more updates for revision

However, the ability of biodiversity-ecosystem functioning (BEF) experiments to accurately represent real-world dynamics is debated [@Wardle2016; @Eisenhauer2016].
Much of the debate centers on the fact that BEF experimental protocols do not allow species gains to offset species losses.
Theoretical work on diversity-\new{ecosystem variability} relationships typically suffers from the same limitation: it recognizes the role of environmental variability in driving population fluctuations which destabilize ecosystems, but ignores the potential for environmental variability to promote species richness and thereby help stabilize ecosystems [Loreau 2010, but see @Chesson2001].

Fluctuating environmental conditions are an important ingredient for stable species coexistence, both in theoretical models [@Chesson2000; @Chesson2004] and in natural communities [@Caceres1997; @Descamps-Julien2005; @Adler2006; @Angert2009].
Such "fluctuation-dependent" coexistence requires that species have unique environmental responses and that environmental conditions vary so that each species in a community experiences favorable and unfavorable conditions, which prevents competitive exclusion [@Chesson2000].
When coexistence is maintained by a fluctuation-dependent mechanism, an increase in environmental variability might lead to an increase in species richness and, consequently, a decrease in ecosystem variability.
However, increasing environmental variability may also increase ecosystem variability by increasing the fluctuations of individual species, regardless of species richness. 
These countervailing effects of environmental variability present an interesting paradox: while we should expect an increase in environmental fluctuations to increase ecosystem variability, this increase might be buffered if fluctuation-dependent coexistence adds new species to the community. 
Such a paradox complicates predictions about how ecosystems will respond to predicted departures from historical ranges of environmental variability.

The opposing effects of environmental variability on ecosystem variability might explain the mixed results from empirical studies on the diversity-\new{ecosystem variability} relationship.
Observational tests of the diversity-\new{ecosystem variability} relationship, which require sampling across natural diversity gradients, have yielded negative [@Hautier2014], neutral [@Valone2003; @Cusson2015], and positive [@Sasaki2011] relationships.
In a meta-analysis of diversity-\new{ecosystem variability} relationships based on observational studies in terrestrial ecosystems, @Jiang2009 found no significant evidence for an effect of species richness on ecosystem variability, perhaps because environmental variability varies across natural diversity gradients, affecting both richness and ecosystem variability.
The idiosyncratic results of these empirical studies contrast with the consistent conclusions from theoretical work that ignores feedbacks between variability and richness.

@@

#### Implementing the Storage Effect
For the storage effect to operate, we need species-specific responses to environmental variability, density-dependent covariance between environmental conditions and competition (*EC* covariance), and subadditive population growth [@Chesson1994; @Chesson2000a]. 
Long-term coexistence is possible when all species can increase when rare.
In the storage effect, rare species increase by escaping the effects of *EC* covariance.
Common species will experience greater than average competition (*C*) in good environment (*E*) years because common species cannot avoid intraspecific competition.
However, rare species can escape intraspecific competition and have the potential to increase rapidly in a good *E* year.
*EC* covariance emerges in our model because dormant-to-live transition rates ($\gamma$) are species-specific and vary through time.
In a high $\gamma$ year for a common species, resource uptake will be above average, while in a high $\gamma$ year for a rare species, resource uptake will be below average, as long as $\gamma$s are not perfectly correlated between species.
Subadditive population growth buffers populations against large population decreases in unfavorable years. 
It is included in our model through a dormant stage with very low death rates, which limits large population declines in bad *E* years.

@@

\begin{figure}[!ht]
  \centering
      \includegraphics[height=5in]{/Users/atredenn/Google_Drive/coexistence-stability_submission/figures/coex_stability_figure1.png}
  \caption{Variability of total community biomass as a function of species richness when coexistence is maintained by the storage effect (A,B) or relative nonlinearity (C,D). Top panels show results from simulations where environmental or resource variance determine the number species that coexist in a community. Bottom panels show results from simulations where environmental or resource variance is fixed at a level that allows coexistence of all four species, but species are removed to manipulate diversity. The left-hand panels represent regional diversity-variability relationships across natural diversity gradients, whereas the right-hand panels represent local diversity-variability relationships. Parameter values, where species are denoted by numeric subscripts: (A) $r_1 = r_2 = r_3 = r_4 = 0.2$, $a_1 = a_2 = a_3 = a_4 = 2$, $b_1 = b_2 = b_3 = b_4 = 2.5$, $\alpha_1 = 0.5, \alpha_2 = 0.49, \alpha_3 = 0.48, \alpha_4 = 0.47$, $\rho_1 = \rho_2 = \rho_3 = \rho_4 = 0$, $\sigma_E =$ variable; (B) $r_1 = r_2 = r_3 = r_4 = 0.2$, $a_1 = a_2 = a_3 = a_4 = 2$, $b_1 = b_2 = b_3 = b_4 = 2.5$, $\alpha_1 = 0.5, \alpha_2 = 0.49, \alpha_3 = 0.48, \alpha_4 = 0.47$, $\rho_1 = \rho_2 = \rho_3 = \rho_4 = -1/3$, $\sigma_E = 4$; (C) $r_1 = 0.2, r_2 = 1, r_3 = 2, r_4 = 5$, $a_1 = 2, a_2 = 5, a_3 = 10, a_4 = 25$, $b_1 = 2.5, b_2 = 20, b_3 = 30, b_4 = 45$, $\alpha_1 = \alpha_2 = \alpha_3 = \alpha_4 = 0.5$, $\rho_1 = \rho_2 = \rho_3 = \rho_4 = 1$, $\sigma(R^+) =$ variable; (D) $r_1 = 0.2, r_2 = 1, r_3 = 2, r_4 = 5$, $a_1 = 2, a_2 = 5, a_3 = 10, a_4 = 25$, $b_1 = 2.5, b_2 = 20, b_3 = 30, b_4 = 45$, $\alpha_1 = \alpha_2 = \alpha_3 = \alpha_4 = 0.5$, $\rho_1 = \rho_2 = \rho_3 = \rho_4 = 1$, $\sigma(R^+) =$ 1.2.}
\end{figure}

>>>>

However, the ability of biodiversity-ecosystem functioning (BEF) experiments to accurately represent real-world dynamics is debated [@Wardle2016; @Eisenhauer2016].
Much of the debate centers on the fact that BEF experimental protocols \new{do not allow species additions from the regional pool to offset species losses in local communities.}
Theoretical work on diversity-\new{ecosystem variability} relationships typically suffers from the same limitation: it recognizes the role of environmental variability in driving population fluctuations which destabilize ecosystems, but ignores the potential for environmental variability to promote species richness and thereby help stabilize ecosystems [Loreau 2010, but see @Chesson2001].

Fluctuating environmental conditions are an important ingredient for stable species coexistence, both in theoretical models [@Chesson2000; @Chesson2004] and in natural communities [@Caceres1997; @Descamps-Julien2005; @Adler2006; @Angert2009].
Such "fluctuation-dependent" coexistence \new{emerges most easily when} species have unique environmental responses and that environmental conditions vary so that each species in a community experiences favorable and unfavorable conditions, which prevents competitive exclusion [@Chesson2000].
When coexistence is maintained by a fluctuation-dependent mechanism, an increase in environmental variability might lead to an increase in species richness and, consequently, a decrease in ecosystem variability.
However, increasing environmental variability may also increase ecosystem variability by increasing the fluctuations of individual species, regardless of species richness. 
These countervailing effects of environmental variability present an interesting paradox: while we should expect an increase in environmental fluctuations to increase ecosystem variability, this increase might be buffered if fluctuation-dependent coexistence adds new species to the community. 
Such a paradox complicates predictions about how ecosystems will respond to predicted departures from historical ranges of environmental variability.

The opposing effects of environmental variability on ecosystem variability might explain the mixed results from \new{observational} studies on the diversity-\new{ecosystem variability} relationship.
Observational tests of the diversity-\new{ecosystem variability} relationship, which require sampling across natural diversity gradients, have yielded negative [@Hautier2014], neutral [@Valone2003; @Cusson2015], and positive [@Sasaki2011] relationships.
In a meta-analysis of diversity-\new{ecosystem variability} relationships based on observational studies in terrestrial ecosystems, @Jiang2009 found no significant evidence for an effect of species richness on ecosystem variability, perhaps because environmental variability varies across natural diversity gradients, affecting both richness and ecosystem variability.
The idiosyncratic results of these observational studies contrast with the consistent conclusions from \new{experimental and theoretical work that ignore, or control, the} feedbacks between variability and richness.

@@

#### Implementing the Storage Effect
For the storage effect to operate, we need species-specific responses to environmental variability, density-dependent covariance between environmental conditions and competition (*EC* covariance), and subadditive population growth [@Chesson1994; @Chesson2000a]. 
Long-term coexistence is possible when all species can increase when rare.
In the storage effect, rare species increase by escaping the effects of *EC* covariance.
Common species will experience greater than average competition (*C*) in good environment (*E*) years because common species cannot avoid intraspecific competition.
However, rare species can escape intraspecific competition and have the potential to increase rapidly in a good *E* year.
*EC* covariance emerges in our model because dormant-to-live transition rates ($\gamma$) are species-specific and vary through time.
In a high $\gamma$ year for a common species, resource uptake will be above average, while in a high $\gamma$ year for a rare species, resource uptake will be below average, as long as $\gamma$s are not perfectly correlated between species.
\new{In other words, resource uptake is above average in a year when an abundant species experiences high germination rates because combined population size will then be above average, but relatively small in a year when a rare species experiences high germination rates because combined population size will be below average.}
Subadditive population growth buffers populations against large population decreases in unfavorable years. 
It is included in our model through a dormant stage with very low death rates, which limits large population declines in bad *E* years.

@@

\begin{figure}[!ht]
  \centering
      \includegraphics[height=5in]{/Users/atredenn/Google_Drive/coexistence-stability_submission/figures/coex_stability_figure1.png}
  \caption{Variability of total community biomass as a function of species richness when coexistence is maintained by the storage effect (A,B) or relative nonlinearity (C,D). Left panels show results from simulations where environmental or resource variance determine the number species that coexist in a community. Right panels show results from simulations where environmental or resource variance is fixed at a level that allows coexistence of all four species, but species are removed to manipulate diversity. The left-hand panels represent regional diversity-variability relationships across natural diversity gradients, whereas the right-hand panels represent local diversity-variability relationships. Parameter values, where species are denoted by numeric subscripts: (A) $r_1 = r_2 = r_3 = r_4 = 0.2$, $a_1 = a_2 = a_3 = a_4 = 2$, $b_1 = b_2 = b_3 = b_4 = 2.5$, $\alpha_1 = 0.5, \alpha_2 = 0.49, \alpha_3 = 0.48, \alpha_4 = 0.47$, $\rho_1 = \rho_2 = \rho_3 = \rho_4 = 0$, $\sigma_E =$ variable; (B) $r_1 = r_2 = r_3 = r_4 = 0.2$, $a_1 = a_2 = a_3 = a_4 = 2$, $b_1 = b_2 = b_3 = b_4 = 2.5$, $\alpha_1 = 0.5, \alpha_2 = 0.49, \alpha_3 = 0.48, \alpha_4 = 0.47$, $\rho_1 = \rho_2 = \rho_3 = \rho_4 = -1/3$, $\sigma_E = 4$; (C) $r_1 = 0.2, r_2 = 1, r_3 = 2, r_4 = 5$, $a_1 = 2, a_2 = 5, a_3 = 10, a_4 = 25$, $b_1 = 2.5, b_2 = 20, b_3 = 30, b_4 = 45$, $\alpha_1 = \alpha_2 = \alpha_3 = \alpha_4 = 0.5$, $\rho_1 = \rho_2 = \rho_3 = \rho_4 = 1$, $\sigma(R^+) =$ variable; (D) $r_1 = 0.2, r_2 = 1, r_3 = 2, r_4 = 5$, $a_1 = 2, a_2 = 5, a_3 = 10, a_4 = 25$, $b_1 = 2.5, b_2 = 20, b_3 = 30, b_4 = 45$, $\alpha_1 = \alpha_2 = \alpha_3 = \alpha_4 = 0.5$, $\rho_1 = \rho_2 = \rho_3 = \rho_4 = 1$, $\sigma(R^+) =$ 1.2.}
\end{figure}
